---
title: "Extracting ICD-10 codes for palliative care estimates in the "
author: "Jan Savinc"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: 
  html_document:
    toc: true
    toc_float: true
editor_options: 
  chunk_output_type: console
bibliography: bibliography.bib
---
  
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Libraries

```{r, warning=FALSE}
library(tidyverse)
library(knitr)
library(pdftools)  # for extracting data from pdf files
```

# Overview

This script loads `appendix B` from the Quebec National Institut report on End of Life Palliative care [@bedard_soins_2006] and parses the ICD-10 codes reported.

# Load document

There was an automatically translated version made previously, however, I haven't been able to extract text from it (it was saved as images, probably?), so I'm using the original, French, text. I've manually trimmed the document to only the pages listing ICD-10 codes in appendix B for faster loading (using Microsoft print as PDF).

```{r}
text_raw <- pdf_text("./QuebecNationalInstitut_ICD10_codes.pdf")
# text_raw <- pdf_text("./QuebecNationalInstitut_ICD10_codes_translated.pdf")  # the translation doesn't load any text - probably the text is saved as imagesd!

# this converts the strings into a table so we can use tidy techniques
text_table <-
  map_dfr(
    .x = text_raw,
    .f = ~str_split(.x, pattern="\\\r\\\n") %>% set_names(., nm="entries")  # split all lines on the \r\n end of line characters, and then also merge the whole thing into a tibble
  ) %>%
  slice(-(1:14)) %>%  # remove first 14 lines, they have to do with ICD-9
  slice(-1) #%>%  # remove the heading saying this is icd-10
```


# Investigating the text

We now have a fairly messy text version of the appendix:

```{r}
head(text_table)
tail(text_table)
```

There is whitespace that was used to align the entries in some hierarchical fashion, and there are unicode characters representing special characters (e.g. `d\u0092` in the text above). What we're interested in is just the ICD codes - the meanings can be inferred automatically later. I could extract the groupings of the codes also but there's little benefit to that and they seem to be following the chapters in ICD-10 anyway!

# Parsing ranges of 3rd digit codes

Some codes are specified at 3rd digit level, and to keep the representation compact, the 3rd level digits are separated by dots, e.g. `A52.0.1.2.3.7`. These need to be interpreted as `A52.0`, `A52.1`, etc.

We can interpret these codes as a prefix `A52` followed by multiple suffixes, listed as dot-separated values. In other words, we split the string at the dots, and then treat the first value as prefix and the rest as suffixes to be concatenated onto the prefix individually.

```{r}
parse_code_range <- function(code_range) {
  if (!str_detect(code_range, pattern="\\..*\\.")) return(code_range)  # return the original input if there aren't at least two dots in the code!
  splits <- str_split(code_range, pattern="\\.", simplify = TRUE)
  prefix <- splits[1]
  suffixes <- splits[2:length(splits)]
  codes_decimal <- paste(prefix, suffixes, sep = ".") 
  codes_csv <- paste0(codes_decimal, collapse = ",")  # concatenate all the separate values into a single string - this makes it easier to check the interim result and allows to separate them later using separate_rows()
}


```


# Extracting the codes from the text

The codes are always specified in the decimal ICD-10 format, e.g. an upper-case letter (A-Z) followed by 2 digits (or more), with some codes specified as ranges as shown above. The codes are always enclosed in parentheses. To extract the codes, then, we'll use a regular expression capturing the open and closing parentheses, capturing the starting letter, two digits, and any more symbols afterwards.

```{r}
crude_extraction <-
  text_table %>%
  filter(str_detect(entries, pattern="\\([A-Z]\\d{2}")) %>%  # keep only rows with the parentheses + code pattern described above
  mutate(
    codes = str_extract(string = entries, pattern = "\\([A-Z]\\d.*\\)"),  # extract the code
    codes = str_remove_all(string = codes, pattern = "\\(|\\)")  # remove parentheses
    )

crude_extraction %>% print(n=10)
```

We now need to parse the codes specified as ranges. I'll also count the nubmer of codes we'd get from it so we can check the results.

What codes were specified as ranges?

```{r}
crude_extraction %>% filter(str_detect(codes, pattern="\\..*\\."))

crude_extraction %>% filter(str_detect(codes, pattern="\\..*\\.")) %>% 
  group_by(entries) %>%
  summarise(codes = codes, code_parsed = parse_code_range(codes), n = length(str_split(code_parsed, pattern = "\\,", simplify = TRUE)))
```

Looks like this works fine, so we can now extract all codes:

```{r}
codes_extracted <-
  crude_extraction %>%
  group_by(codes) %>%
  summarise(code_parsed = parse_code_range(codes), .groups="drop") %>%
  separate_rows(code_parsed, sep=",")

codes_final <- 
  codes_extracted %>% 
  select(code_decimal = code_parsed) %>%  # keep only the parsed code
  distinct  # remove duplicates
```

Finally, for reference, we'll print all of the codes:

```{r}
codes_extracted %>% print(n=nrow(.))
```


# Categorising trajectories

The codes can be categorised as one of three possible illness trajectories:

* Trajectory 1: short period of evident decline, typically cancer - ‘tumours’
* Trajectory 2: long-term limitations with intermittent serious episodes
* Trajectory 3: prolonged dwindling

The article [@bedard_soins_2006] specifies the ICD-10 codes in groups corresponding to the ICD-10 chapters; the Trajectories 1 & 3 are specified as one or two chapters, whereas Trajectory 2 includes all other chapters, so that's the order of coding I'll do.

This is what the article says:

**Trajectory 1**

> La trajectoire I correspond à une évolution progressive et à phase terminale habituellement relativement claire. Selon Lunney et al. (13), Murray et al. (14) et Murtagh et al. (15), les décès par cancer sont typiques de la trajectoire I. Nous avons donc assigné la trajectoire I à tous les décès par « Tumeurs » (voir codes CIM-9 et CIM-10 à l’annexe B).

This trajectory includes tumours, which in in the appendix are defined as all C__ codes and D__ codes up to including D48.


**Trajectory 2**

> La trajectoire II se définit par un déclin graduel, ponctué par des épisodes de
détérioration aigus et certains moments de récupération, avec une mort parfois
soudaine et inattendue. Selon Lunney et al. (13), Murray et al. (14) et Murtagh et al.
(15), la trajectoire II décrit la trajectoire des patients qui décèdent de défaillance du
système cardio-pulmonaire. Nous avons donc assigné la trajectoire II aux décès par
« Maladies de l’appareil circulatoire » et de « Maladies de l’appareil respiratoire »
(voir codes CIM-9 et CIM-10 à l’annexe B). À la lumière des résultats, il est apparu
que la trajectoire II semblait décrire également l’évolution de plusieurs autres
maladies chroniques qui se sont avérées fatales notamment de « Maladies
endocriniennes, de la nutrition et du métabolisme », de « Maladies de l’appareil
digestif », de « Maladies de l’appareil génito-urinaire », de « Maladies de l’appareil
ostéo-articulaire, des muscles et du tissu conjonctif, maladies de la peau », de
« Maladies infectieuses et parasitaires », de « Malformations congénitales et
anomalies chromosomiques » et de «Maladies du sang et des organes
hématopoïétiques et certains troubles du système immunitaire » (voir codes CIM-9 et
CIM-10 à l’annexe B).

This trajectory includes:

* Disorders of the circulatory system,
* Disorders of the pulmonary system,
* Endocrine disorders & disorders of nutrition and metabolism,
* Digestive organs disorders,
* Genito-urinary disroders,
* Skeleto-muscular disorders,,
* Infectious & parasitic diseaeses,
* Congenital & chromosomal disorders,
* Blood disorders & certain immune system disorders

This also includes the R__ codes for signs & symptoms!

Practically, this includes all codes not covered by Trajectories 1 & 3!


**Trajectory 3**

> La trajectoire III se définit par un déclin graduel et prolongé, typique des personnes
âgées et fragiles ou des personnes avec démence selon Lunney et al. (13), Murray
et al. (14) et Murtagh et al. (15). Nous avons donc assigné cette trajectoire aux «
Troubles mentaux » (CIM-9), « Troubles mentaux et du comportement » (CIM-10),
« Maladies du système nerveux et des organes des sens » (CIM-9) et « Maladies du
système nerveux » (CIM-10) (voir codes CIM-9 et CIM-10 à l’annexe B).

This trajectory includes:

* mental & behavioural disorders (F00-F06)
* nervous system disorders (G03-G99)

In this list of codes, this captures all F__ and G__ codes!

```{r}
codes_final <-
  codes_final %>%
  mutate(
    trajectory = case_when(
      str_detect(code_decimal, pattern=paste0("^C|", paste0("^",paste0("D",str_pad(string = 0:48, width = 2, pad = "0")),collapse="|"))) ~ "Trajectory 1: short period of evident decline",
      str_detect(code_decimal, pattern="^F|^G") ~ "Trajectory 3: prolonged dwindling",
       TRUE ~ "Trajectory 2: long-term limitations with intermittent serious episodes"
    )
  )
```


These will be saved as `.csv`:

```{r}
write_csv(x = codes_final, path = "./outputs/quebec_palliative_icd10_codes.csv")
```



# References