---
title: "Producing tables for Migrations paper from safe haven exported data"
author: "Jan Savinc"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: 
  html_document:
  toc: true
toc_float: true
editor_options: 
  chunk_output_type: console
---
  
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Libraries

```{r, warning=FALSE}
library(tidyverse)
library(readxl)  # for reading excel files
library(gt)  # for formating tables
library(flextable)  # for formating tables
library(knitr)
library(viridis)  # for colour scheme
```

# Overview

This script loads the file `MigrationConf.xlsx`, and extracts data from the tables therein which we can then format in various ways for the paper!

# Load exported data

```{r}
exported_file <- "./raw/MigrationConf.xlsx"
available_sheets <- excel_sheets(exported_file)
```

Loading the data won't be straightforward - there are multiple tables in a sheet, so we'll need to specify where those tables are.

## `P` sheet

The `P` sheet contains descriptives - 2-way breakdowns of Urban-rural (6 category) indicator and other variables. 

```{r}
## obtained by copying from the excel table using datapasta::vector_paste()
urban_rural_6_categories <- c("Large Urban Areas", "Other Urban Areas", "Accessible Small Towns", "Remote Small Towns", "Accessible Rural", "Remote Rural")

odds <- function(nums) {  # return odd numbers in range
  nums[(nums %% 2) != 0]
}
  
urban_rural_breakdown <- 
  read_excel(
    path = exported_file,
    sheet = "P",
    range = "O31:U31",
    col_names = c(urban_rural_6_categories, "total")
  ) %>% 
  pivot_longer(cols = everything(), names_to = "urban_rural_6_categories", values_to = "n")

total_n <- urban_rural_breakdown$n[urban_rural_breakdown$urban_rural_6_categories=="total"]

variables_in_p <- c("sex","age_at_death","marital_status_at_death","country_of_birth")

data_in_d <- map2(
  .x = variables_in_p,
  .y = c("N7:U10", "N11:U16", "N17:U24", "N25:U30"),  # manually found ranges of data in the file
  .f = ~read_excel(
          path = exported_file,
          sheet = "P",
          range = .y,  # pass the range for each variable
          col_names = c(.x, urban_rural_6_categories, "total")  # construct column names from variable of interest & urban-rural breakdown
        ) %>%
        slice(odds(1:nrow(.)))  # keep the odd numbered rows only - the even-numbered rows contain percentages of total and we can calculate these ourselves
  ) %>% set_names(variables_in_p)
```

### Missing value summary

```{r}
missing_value_summary <-
  read_excel(
    path = exported_file,
    sheet = "P",
    range = "C12:H15",  # skipping the I column with total percentage (which is 100% anyway)
    col_names = c("variable", "n_valid", "percent_valid", "n_missing", "percent_missing", "n_total")
  )
```

## `D` sheet

The `D` sheet contains a 3-way breakdown of Urban-rural (6 category) indicator, country of birth, and living alone.

It looks like there were more breakdowns planned but not completed.

```{r}
country_of_birth_categories <- c("Scotland", "Rest of UK", "Outside UK", "All CoB")  # obtained by copying from the excel table using datapasta::vector_paste()
levels_urban_rural_6_categories <-  # levels of the variable so they appear in the right order
  c(
    "Large Urban Areas",
    "Other Urban Areas",
    "Accessible Small Towns",
    "Remote Small Towns",
    "Accessible Rural",
    "Remote Rural"
  )

living_alone_table_header <- c(  # compile the column names to match what's in the spreadsheet, with a prefix that we'll use to split the columns up later
  "urban_rural_6_categories",
  paste0("alone_", country_of_birth_categories),
  paste0("total_", country_of_birth_categories)
  )

living_alone_by_cob_and_urban_rural <-
  read_excel(
    path = exported_file,
    sheet = "D",
    range = "C14:K25",
    col_names = living_alone_table_header
  ) %>%
  slice(odds(1:nrow(.))) %>%
  pivot_longer(cols = -urban_rural_6_categories, names_to = c("living_alone","country_of_birth"), names_pattern = "^(alone|total)\\_(.*)$", values_to = "n") %>%  # split the columns containing multiple variables (living alone or not & CoB)
  mutate(
    country_of_birth = factor(country_of_birth, levels = country_of_birth_categories),
    urban_rural_6_categories = factor(urban_rural_6_categories, levels = levels_urban_rural_6_categories)
  )
  
living_alone_by_cob <-
  read_excel(
    path = exported_file,
    sheet = "D",
    range = "D26:K26",
    col_names = living_alone_table_header[-1],  # remove the first variable in the constructed column names from above since we're not using the urban rural column here
  ) %>%
  pivot_longer(cols = everything(), names_to = c("living_alone","country_of_birth"), names_pattern = "^(alone|total)\\_(.*)$", values_to = "n")  # split the columns containing multiple variables (living alone or not & CoB)

## test that we get the right numbers
living_alone_by_cob_and_urban_rural %>%
  group_by(living_alone, country_of_birth) %>%
  summarise(n_sum = sum(n), .groups="drop") %>%
  left_join(living_alone_by_cob)
```

## `M1` to `M3`

These are logistic models:

* M1 is a stratified regression (stratified on 3 urban-rural categories, meaning three separate models) of living alone on the other variables: time between census and death, marital status, age at death, country of birth, tenure (housing), and sex
* M2 is the same model as M1 but the dependent variable is dying in an institution
* M3 is a regression of living alone on all variables in the entire sample (unstratified), including the 3 categories of urban-rurality as an independent variable

As far as I can tell, for the purposes of the paper, M1 and M2 were used, and only the Rural (remote and accessible) stratum was considered (interpreted as the destination of migrations).

M1 also contains the frequency tables for the variables broken down by the 3-way urban-rural indicator, as well as a total in a column to the right.

There are univariate models in the bottom of each sheet.

### M1

```{r}
regression_columns <- c("b", "se", "Wald", "df", "p_value", "or")  # copied from excel & adjusted for more r-like labelling

m1_regression_variables <- c(
  "country_of_birth",
  "age_at_death",
  "sex_at_death",
  "housing_tenure",
  "time_to_death",
  "marital_status"
)

m1_univariate_regression_ranges <- c(
  "C125:I128", # CoB  - this is lsited in the first step of the regressio nrather than unadjusted regressions later!
  "C285:I288",  # age
  "C300:I301",  # sex
  "C314:I317",  # housing tenure
  "C330:I337",  # time to death
  "C348:I352"  # marital status
)


m1_remote_univariate_regressions <-
  map2(
    .x = m1_regression_variables,
    .y = m1_univariate_regression_ranges,
    .f = 
      ~read_excel(
      path = exported_file,
      sheet = "M1",
      range = .y,
      col_names = c(.x, regression_columns),  # remove the first variable in the constructed column names from above since we're not using the urban rural column here
      col_types = c("text",rep("numeric",times=6))
    )
  ) %>% set_names(m1_regression_variables)

## The country of birth value labels are missing so we need to change them manually!
regression_country_of_birth_categories <- country_of_birth_categories[-4]
regression_country_of_birth_categories[1] <- "Scotland (ref: Scotland)"
m1_remote_univariate_regressions$country_of_birth$country_of_birth[1:3] <- regression_country_of_birth_categories

## The sex variable is binary and has only one entry - to make it have the same format as the others we add another row for the reference column
m1_remote_univariate_regressions$sex_at_death <-
  m1_remote_univariate_regressions$sex_at_death %>%
  add_case(sex_at_death = m1_remote_univariate_regressions$sex_at_death$sex_at_death[1], .before=1)
m1_remote_univariate_regressions$sex_at_death$sex_at_death[2] <- "male"
  

m1_multiple_regression_ranges <- c(
  "C239:I241",  # CoB
  "C242:I244",  # age
  "C245:I245",  # sex 
  "C246:I248",  # housing tenure
  "C249:I255",  # time to death
  "C256:I259",  # marital status
  "C260:I260"  #  constant
)

m1_remote_multiple_regressions <-
  map2(
    .x = c(m1_regression_variables,"constant"),
    .y = m1_multiple_regression_ranges,
    .f = 
      ~read_excel(
      path = exported_file,
      sheet = "M1",
      range = .y,
      col_names = c(.x, regression_columns),  # remove the first variable in the constructed column names from above since we're not using the urban rural column here
      col_types = c("text",rep("numeric",times=6))
    )
  ) %>% set_names(c(m1_regression_variables,"constant"))

## the variable labels are missing in the multiple regression, but they can be copied over from the univariate regressions!
m1_remote_multiple_regressions$country_of_birth$country_of_birth <- m1_remote_univariate_regressions$country_of_birth$country_of_birth[1:3]
m1_remote_multiple_regressions$age_at_death$age_at_death <- m1_remote_univariate_regressions$age_at_death$age_at_death[1:3]
m1_remote_multiple_regressions$sex_at_death <-  # we'll create an extra row for the sex variable which is binary & didn't have a separate reference entry like all the others
  m1_remote_multiple_regressions$sex_at_death %>%
  add_case(sex_at_death=NA, .before = 1)
m1_remote_multiple_regressions$sex_at_death$sex_at_death <- m1_remote_univariate_regressions$sex_at_death$sex_at_death[1:2]
m1_remote_multiple_regressions$housing_tenure$housing_tenure <- m1_remote_univariate_regressions$housing_tenure$housing_tenure[1:3]
m1_remote_multiple_regressions$time_to_death$time_to_death <- m1_remote_univariate_regressions$time_to_death$time_to_death[1:7]
m1_remote_multiple_regressions$marital_status$marital_status <- m1_remote_univariate_regressions$marital_status$marital_status[1:4]
```


### M1 frequency tables

The frequency table is in the `M1` sheet under the Categorical Variable Codings heading, range B53:D74, L53:N74, V53:X74 for the urban-rural strata, respectively, and AF53:AH74 for the entire sample.


```{r}
urban_rural_strata <- c("Rural","Small towns & other urban","Large urban")  # the order they appear in the exported file

m1_frequency_table_ranges <- c(
  "B53:D74",  # rural
  "L53:N74",  # small towns
  "V53:X74",  # large urban
  "AF53:AH74"  # entire sample
)

m1_frequency_tables <-
  map2(
    .x = m1_frequency_table_ranges,
    .y = c(urban_rural_strata, "Entire sample"),  # strata with the entire sample added
    .f = function(.x, .y) {
      proportion_var_name = paste0("prop_",janitor::make_clean_names(.y))
      num_var_name = paste0("n_",janitor::make_clean_names(.y))
      read_excel(
          path = exported_file,
          sheet = "M1",
          range = .x,
          col_names = c("variable","levels",num_var_name),
          col_types = c("text","text","numeric")
      ) %>%
      fill(variable, .direction = "down") %>%  # complete the variable names down the line
      group_by(variable) %>%
      mutate(  # define a new proportion variable, calculated from the sum of the num_ for all levels of a variable
        !!rlang::sym(proportion_var_name):= scales::percent(!!rlang::sym(num_var_name) / sum(!!rlang::sym(num_var_name)), accuracy = 0.1)
      ) %>%
      ungroup
    }
  ) %>% set_names(janitor::make_clean_names(c(urban_rural_strata, "Entire sample")))  # name list items appropriately

m1_merged_frequency_table <-
  m1_frequency_tables %>%
  reduce(.x = , .f = function(x,y) full_join(x,y, by=c("variable","levels")))

## check that the entire sample total actually matches the strata totals
stopifnot(
  m1_merged_frequency_table$n_entire_sample == m1_merged_frequency_table$n_rural + m1_merged_frequency_table$n_large_urban + m1_merged_frequency_table$n_small_towns_other_urban
)
```


### M2

```{r}
m2_regression_variables <- c(
  "country_of_birth",
  "age_at_death",
  "sex_at_death",
  "housing_tenure",
  "time_to_death",
  "living_alone",
  "marital_status"
)

m2_univariate_regression_ranges <- c(
  "C117:I120", # CoB  - this is lsited in the first step of the regressio nrather than unadjusted regressions later!
  "C322:I325",  # age
  "C336:I337",  # sex
  "C348:I351",  # housing tenure
  "C362:I369",  # time to death
  "C380:I381",  # living alone
  "C392:I396"  # marital status
)


m2_remote_univariate_regressions <-
  map2(
    .x = m2_regression_variables,
    .y = m2_univariate_regression_ranges,
    .f = 
      ~read_excel(
      path = exported_file,
      sheet = "M2",
      range = .y,
      col_names = c(.x, regression_columns),  # remove the first variable in the constructed column names from above since we're not using the urban rural column here
      col_types = c("text",rep("numeric",times=6))
    )
  ) %>% set_names(m2_regression_variables)

## living alone and sex are binary, for both we should set a reference level so they are in the same format as the other variables
m2_remote_univariate_regressions$sex_at_death <-
  m2_remote_univariate_regressions$sex_at_death %>%
  add_case(sex_at_death = m2_remote_univariate_regressions$sex_at_death$sex_at_death[1], .before=1)
  
m2_remote_univariate_regressions$living_alone <-
  m2_remote_univariate_regressions$living_alone %>%
  add_case(living_alone = m2_remote_univariate_regressions$living_alone$living_alone[1], .before=1)
m2_remote_univariate_regressions$living_alone$living_alone[1] <- "alone (ref: alone)"
m2_remote_univariate_regressions$living_alone$living_alone[2] <- "with others"

## all value labels are missing here so we can copy them from M1, except for living alone
for (variable in names(m1_remote_univariate_regressions)) {
  m2_remote_univariate_regressions[[variable]][[variable]] <- m1_remote_univariate_regressions[[variable]][[variable]]
}

m2_multiple_regression_ranges <- c(
  "C277:I279",  # CoB
  "C280:I282",  # age
  "C283:I283",  # sex 
  "C284:I286",  # housing tenure
  "C287:I293",  # time to death
  "C294:I294",  # living alone
  "C295:I298",  # living alone
  "C299:I299"  #  constant
)

m2_remote_multiple_regressions <-
  map2(
    .x = c(m2_regression_variables,"constant"),  # we include the constant term here also
    .y = m2_multiple_regression_ranges,
    .f = 
      ~read_excel(
      path = exported_file,
      sheet = "M2",
      range = .y,
      col_names = c(.x, regression_columns),  # remove the first variable in the constructed column names from above since we're not using the urban rural column here
      col_types = c("text",rep("numeric",times=6))
    )
  ) %>% set_names(c(m2_regression_variables,"constant"))


## living alone and sex are binary, for both we should set a reference level so they are in the same format as the other variables
m2_remote_multiple_regressions$sex_at_death <-
  m2_remote_multiple_regressions$sex_at_death %>%
  add_case(sex_at_death = m2_remote_multiple_regressions$sex_at_death$sex_at_death[1], .before=1)
  
m2_remote_multiple_regressions$living_alone <-
  m2_remote_multiple_regressions$living_alone %>%
  add_case(living_alone = m2_remote_univariate_regressions$living_alone$living_alone[1], .before=1)

## the variable labels are missing in the multiple regression, but they can be copied over from the univariate regressions!
for (variable in names(m2_remote_univariate_regressions)) {
  number_entries <- length(m2_remote_multiple_regressions[[variable]][[variable]])
  m2_remote_multiple_regressions[[variable]][[variable]] <-
    m2_remote_univariate_regressions[[variable]][[variable]][1:number_entries]
}
```


### M2 frequency table

We can extract a few more pieces of descriptive summary from the M2 model, namely the breakdown of the `living alone` and `died at home or institution` variables. The strate are the same as for M1, and so the breakdown of individuals by all expalantory variables is also the same.

The `living alone` variable is listed in the the Categorical Variable Codings heading, range B65:D66, L65:N66, V65:X66 for the urban-rural strata, and we'll calculate the entire sample breakdown ourselves. The `died at home/institution` variable can be extracted from the observed/predicted matrices exported by SPSS, in ranges C75:F76, M75:P76, W75:Z76.



```{r}
m2_living_alone_ranges <- c(
  "B65:D66",  # rural
  "L65:N66",  # small towns
  "V65:X66"  # large urban
)

m2_living_alone <-
  map2(
    .x = m2_living_alone_ranges,
    .y = urban_rural_strata,  # strata with the entire sample added
    .f = function(.x, .y) {
      proportion_var_name = paste0("prop_",janitor::make_clean_names(.y))
      num_var_name = paste0("n_",janitor::make_clean_names(.y))
      read_excel(
          path = exported_file,
          sheet = "M2",
          range = .x,
          col_names = c("variable","levels",num_var_name),
          col_types = c("text","text","numeric")
      ) %>%
      fill(variable, .direction = "down") %>%  # complete the variable names down the line
      group_by(variable) %>%
      mutate(  # define a new proportion variable, calculated from the sum of the num_ for all levels of a variable
        !!rlang::sym(proportion_var_name):= scales::percent(!!rlang::sym(num_var_name) / sum(!!rlang::sym(num_var_name)), accuracy = 0.1)
      ) %>%
      ungroup
    }
  ) %>% set_names(janitor::make_clean_names(urban_rural_strata))  # name list items appropriately

m2_merged_living_alone <-
  m2_living_alone %>%
  reduce(.x = , .f = function(x,y) full_join(x,y, by=c("variable","levels"))) %>%
  mutate(
    n_entire_sample = rowSums(select(., dplyr::starts_with("n_"))),
    prop_entire_sample = scales::percent(n_entire_sample / sum(n_entire_sample), accuracy = 0.1)
    )


m2_place_of_death_ranges <- c(
  "C75:F76",  # rural
  "M75:P76",  # small town + other urban
  "W75:Z76"  # large urban
)


m2_place_of_death <-
  map2(
    .x = m2_place_of_death_ranges,
    .y = urban_rural_strata,  # strata with the entire sample added
    .f = function(.x, .y) {
      proportion_var_name = paste0("prop_",janitor::make_clean_names(.y))
      num_var_name = paste0("n_",janitor::make_clean_names(.y))
      read_excel(
          path = exported_file,
          sheet = "M2",
          range = .x,
          col_names = c("variable","levels","remove_this",num_var_name),
          col_types = c("text","numeric","numeric","numeric")
      ) %>%
      select(-remove_this) %>%  # remove unneeded column
      fill(variable, .direction = "down") %>%  # complete the variable names down the line
      group_by(variable) %>%
      mutate(  # define a new proportion variable, calculated from the sum of the num_ for all levels of a variable
        !!rlang::sym(proportion_var_name):= scales::percent(!!rlang::sym(num_var_name) / sum(!!rlang::sym(num_var_name)), accuracy = 0.1)
      ) %>%
      ungroup
    }
  ) %>% set_names(janitor::make_clean_names(urban_rural_strata))  # name list items appropriately


m2_merged_place_of_death <-
  m2_place_of_death %>%
  reduce(.x = , .f = function(x,y) full_join(x,y, by=c("variable","levels"))) %>%
  mutate(  # compute entire sample
    n_entire_sample = rowSums(select(., dplyr::starts_with("n_"))),
    prop_entire_sample = scales::percent(n_entire_sample / sum(n_entire_sample), accuracy = 0.1),
    levels = if_else(levels==1, "Institution", "Home")
    )
```


### M3

M3 is a regression on the entire sample of individuals living alone

TODO!

# Notes on data

* Individuals with missing marital status were excluded due to very small N;
* Logistic models used complete data only


# Cleaning up variable names & levels

```{r}
#   tibble(
#     raw = unique(c(m1_merged_frequency_table$variable, m1_regression_variables, m2_regression_variables, m2_merged_living_alone$variable, m2_merged_place_of_death$variable, unique(str_remove_all(m1_remote_multivariate_regression$variable, pattern="\\(\\d\\)"))))
#   )
# datapasta::tribble_paste(clean_variable_names)
clean_variable_names <-
  tibble::tribble(
                                                                 ~raw, ~clean,
                "Time between Census enumeration and death in groups", "Time between Census enumeration and death",
                                                      "time_to_death", "Time between Census enumeration and death",
                                                       "survival_grp", "Time between Census enumeration and death",
                                                   "MARITAL_STATUS_d", "Marital status",
                                                     "marital_status", "Marital status",
                                  "age (at death) in ten year groups", "Age",
                                                       "age_at_death", "Age",
                                                           "age_d10y", "Age",
"Country of Birth from death records (Scotland, rest of UK, elswhere)", "Country of birth",
                                                   "country_of_birth", "Country of birth",
                                                         "CoB_d_3cat", "Country of birth",
                                                        "tenure_3cat", "Tenure type",
                                                     "housing_tenure", "Tenure type",
                                                        "tenure_3cat", "Tenure type",
                                                       "sex at death", "Sex",
                                                       "sex_at_death", "Sex",
                                                              "SEX_d", "Sex",
                                                       "living_alone", "Living alone",
           "Dummy of living alone or with others/in care institution", "Living alone",
                "Died in institution (hospital, care home, hospice)?", "Place of death",
                                              "Urban-rural indicator", "Urban-rural indicator"
     ) %>% distinct
```


# Output directory

```{r}
if (!dir.exists("./outputs_migrations/")) dir.create("./outputs_migrations/")
```



# Processed frequency tables

```{r}
frequency_tables <- list()

urban_rural_frequency <-
  m1_merged_frequency_table %>% filter(variable=="sex at death") %>%
  summarise_if(is.numeric, sum) %>% 
  select(-n_entire_sample) %>%
  pivot_longer(cols = everything(), names_to = "levels", values_to = "n_entire_sample") %>%
  mutate(
    variable = "Urban-rural indicator", 
    levels = urban_rural_strata,
    prop_entire_sample = scales::percent(n_entire_sample / sum(n_entire_sample), accuracy = 0.1)
      )

frequency_tables$descriptives <-
  bind_rows(
    m1_merged_frequency_table,
    m2_merged_living_alone,
    m2_merged_place_of_death,
    urban_rural_frequency
  ) %>%
  left_join(
    clean_variable_names, by = c("variable"="raw")
  ) %>%
  distinct %>%
  mutate(
    variable = clean_variable_names$clean[match(x = variable, table = clean_variable_names$raw)],
    levels = str_to_sentence(levels)
  )

## save frequency table
write_csv(frequency_tables$descriptives, path = "./outputs_migrations/migrations_descriptives_all_variable_frequencies_by_urban_rural_3_cat.csv")
```


# Processed regression tables

```{r}
regression_tables <- list()

regression_tables$m1_remote_unadjusted_odds <-
  map2_df(
    .x = m1_remote_univariate_regressions,
    .y = names(m1_remote_univariate_regressions),
    .f = function(x,y) {
      tbl <- select(x, 1, or) %>%
      mutate(variable = y) %>%
      rename(level=1) %>%
      filter(level!="Constant") %>%
      relocate(variable, level, or)
      
      tbl$level[1] <- str_replace(string = tbl$level[1], pattern=".*ref\\:\\s(.*)\\)", replacement="Reference: \\1")  # rename the reference level
      # tbl$or[1] <- 1L  # set reference OR to 1
      tbl$or[1] <- NA  # set reference OR to NA
      return(tbl)
    } 
  )

regression_tables$m1_remote_adjusted_odds <-
  map2_df(
    .x = m1_remote_multiple_regressions[names(m1_remote_multiple_regressions)!="constant"],  # don't incldue the Constant term
    .y = names(m1_remote_univariate_regressions[names(m1_remote_multiple_regressions)!="constant"]),  # see above comment
    .f = function(x,y) {
      tbl <- select(x, 1, or) %>%
      mutate(variable = y) %>%
      rename(level=1) %>%
      filter(level!="Constant") %>%
      relocate(variable, level, or)
      
      tbl$level[1] <- str_replace(string = tbl$level[1], pattern=".*ref\\:\\s(.*)\\)", replacement="Reference: \\1")  # rename the reference level
      # tbl$or[1] <- 1L  # set reference OR to 1
      tbl$or[1] <- NA  # set reference OR to NA
      return(tbl)
    } 
  )

regression_tables$m1_odds <-
  full_join(
    x = regression_tables$m1_remote_unadjusted_odds %>% rename(unadjusted_or = or),
    y = regression_tables$m1_remote_adjusted_odds %>% rename(adjusted_or = or),
    by = c("variable", "level")
  ) %>%
  mutate(
    variable = clean_variable_names$clean[match(x = variable, table = clean_variable_names$raw)],
    level = str_to_sentence(level)
  )


regression_tables$m2_remote_unadjusted_odds <-
  map2_df(
    .x = m2_remote_univariate_regressions,
    .y = names(m2_remote_univariate_regressions),
    .f = function(x,y) {
      tbl <- select(x, 1, or) %>%
      mutate(variable = y) %>%
      rename(level=1) %>%
      filter(level!="Constant") %>%
      relocate(variable, level, or)
      
      tbl$level[1] <- str_replace(string = tbl$level[1], pattern=".*ref\\:\\s(.*)\\)", replacement="Reference: \\1")  # rename the reference level
      # tbl$or[1] <- 1L  # set reference OR to 1
      tbl$or[1] <- NA  # set reference OR to NA
      return(tbl)
    } 
  )

regression_tables$m2_remote_adjusted_odds <-
  map2_df(
    .x = m2_remote_multiple_regressions[names(m2_remote_multiple_regressions)!="constant"],  # don't incldue the Constant term
    .y = names(m2_remote_univariate_regressions[names(m2_remote_multiple_regressions)!="constant"]),  # see above comment
    .f = function(x,y) {
      tbl <- select(x, 1, or) %>%
      mutate(variable = y) %>%
      rename(level=1) %>%
      filter(level!="Constant") %>%
      relocate(variable, level, or)
      
      tbl$level[1] <- str_replace(string = tbl$level[1], pattern=".*ref\\:\\s(.*)\\)", replacement="Reference: \\1")  # rename the reference level
      # tbl$or[1] <- 1L  # set reference OR to 1
      tbl$or[1] <- NA  # set reference OR to NA
      return(tbl)
    } 
  )

regression_tables$m2_odds <-
  full_join(
    x = regression_tables$m2_remote_unadjusted_odds %>% rename(unadjusted_or = or),
    y = regression_tables$m2_remote_adjusted_odds %>% rename(adjusted_or = or),
    by = c("variable", "level")
  ) %>%
  mutate(
    variable = clean_variable_names$clean[match(x = variable, table = clean_variable_names$raw)],
    level = str_to_sentence(level)
  )


# TODO: look at formatting regression_tables with `gt`

## save frequency table
write_csv(regression_tables$m1_odds, path = "./outputs_migrations/migrations_logistic_regression_living_alone_rural_stratum.csv")

## save frequency table
write_csv(regression_tables$m2_odds, path = "./outputs_migrations/migrations_logistic_regression_death_in_institution_rural_stratum.csv")
```

# Processed figures

```{r}
figures <- list()
```

## Proportion living alone by CoB & Urban-rural (6-fold) indicator

```{r}
## notes: scotland is by far the most frequent category, so displaying average of all CoB is moot - it's almsot exactly the same as Scotland
figures$alone_by_country_of_birth_and_urban_rural <-
  ggplot(
    data = living_alone_by_cob_and_urban_rural %>% pivot_wider(names_from=living_alone, values_from=n) %>% mutate(proportion=alone/total) %>% filter(country_of_birth!="All CoB")
  ) +
    geom_col(aes(x=urban_rural_6_categories, y=proportion, group=country_of_birth, fill=country_of_birth), position = position_dodge()) +
    scale_fill_viridis(discrete = TRUE) +
    scale_x_discrete(labels = function(x) str_wrap(x, width = 10)) +
    scale_y_continuous(labels = scales::label_percent(accuracy = 1L)) +
    labs(
      x = "Urban-rural indicator (6-fold, 2012 data)",
      y = "Proportion individuals living alone (%)", fill = "Country of birth"
    ) +
    theme_minimal() +
    # theme(axis.text.x = element_text(angle = -45, hjust = 0)) +  # looks good enough with str_wrap, no need for rotated text
    NULL

figures$alone_by_country_of_birth_and_urban_rural_grey <-  # greyscale version of above
  figures$alone_by_country_of_birth_and_urban_rural +
  scale_fill_grey()

ggsave(figures$alone_by_country_of_birth_and_urban_rural, filename = "./outputs_migrations/proportion_living_alone_by_cob_and_urban_rural_6cat.pdf")
ggsave(figures$alone_by_country_of_birth_and_urban_rural, filename = "./outputs_migrations/proportion_living_alone_by_cob_and_urban_rural_6cat.svg")
ggsave(figures$alone_by_country_of_birth_and_urban_rural_grey, filename = "./outputs_migrations/proportion_living_alone_by_cob_and_urban_rural_6cat_grey.pdf")
ggsave(figures$alone_by_country_of_birth_and_urban_rural_grey, filename = "./outputs_migrations/proportion_living_alone_by_cob_and_urban_rural_6cat_grey.svg")
```


## Proportion died in institution

TODO: data for place of death cross-tabulation not exported from safe haven yet