---
title: "Producing tables for Migrations paper from safe haven exported data"
author: "Jan Savinc"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: 
  html_document:
  toc: true
toc_float: true
editor_options: 
  chunk_output_type: console
---
  
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Libraries

```{r, warning=FALSE}
library(tidyverse)
library(readxl)  # for reading excel files
library(gt)  # for formating tables
library(flextable)  # for formating tables
library(knitr)
```

# Overview

This script loads the file `MigrationConf.xlsx`, and extracts data from the tables therein which we can then format in various ways for the paper!

# Load exported data

```{r}
exported_file <- "./raw/MigrationConf.xlsx"
available_sheets <- excel_sheets(exported_file)
```

Loading the data won't be straightforward - there are multiple tables in a sheet, so we'll need to specify where those tables are.

## `P` sheet

The `P` sheet contains descriptives - 2-way breakdowns of Urban-rural (6 category) indicator and other variables. 

```{r}
## obtained by copying from the excel table using datapasta::vector_paste()
urban_rural_6_categories <- c("Large Urban Areas", "Other Urban Areas", "Accessible Small Towns", "Remote Small Towns", "Accessible Rural", "Remote Rural")

odds <- function(nums) {  # return odd numbers in range
  nums[(nums %% 2) != 0]
}
  
urban_rural_breakdown <- 
  read_excel(
    path = exported_file,
    sheet = "P",
    range = "O31:U31",
    col_names = c(urban_rural_6_categories, "total")
  ) %>% 
  pivot_longer(cols = everything(), names_to = "urban_rural_6_categories", values_to = "n")

total_n <- urban_rural_breakdown$n[urban_rural_breakdown$urban_rural_6_categories=="total"]

variables_in_p <- c("sex","age_at_death","marital_status_at_death","country_of_birth")

data_in_d <- map2(
  .x = variables_in_p,
  .y = c("N7:U10", "N11:U16", "N17:U24", "N25:U30"),  # manually found ranges of data in the file
  .f = ~read_excel(
          path = exported_file,
          sheet = "P",
          range = .y,  # pass the range for each variable
          col_names = c(.x, urban_rural_6_categories, "total")  # construct column names from variable of interest & urban-rural breakdown
        ) %>%
        slice(odds(1:nrow(.)))  # keep the odd numbered rows only - the even-numbered rows contain percentages of total and we can calculate these ourselves
  ) %>% set_names(variables_in_p)
```

### Missing value summary

```{r}
missing_value_summary <-
  read_excel(
    path = exported_file,
    sheet = "P",
    range = "C12:H15",  # skipping the I column with total percentage (which is 100% anyway)
    col_names = c("variable", "n_valid", "percent_valid", "n_missing", "percent_missing", "n_total")
  )
```

## `D` sheet

The `D` sheet contains a 3-way breakdown of Urban-rural (6 category) indicator, country of birth, and living alone.

It looks like there were more breakdowns planned but not completed.

```{r}
country_of_birth_categories <- c("Scotland", "Rest of UK", "Outside UK", "All CoB")  # obtained by copying from the excel table using datapasta::vector_paste()


living_alone_table_header <- c(  # compile the column names to match what's in the spreadsheet, with a prefix that we'll use to split the columns up later
  "urban_rural_6_categories",
  paste0("alone_", country_of_birth_categories),
  paste0("total_", country_of_birth_categories)
  )

living_alone_by_cob_and_urban_rural <-
  read_excel(
    path = exported_file,
    sheet = "D",
    range = "C14:K25",
    col_names = living_alone_table_header
  ) %>%
  slice(odds(1:nrow(.))) %>%
  pivot_longer(cols = -urban_rural_6_categories, names_to = c("living_alone","country_of_birth"), names_pattern = "^(alone|total)\\_(.*)$", values_to = "n")  # split the columns containing multiple variables (living alone or not & CoB)

living_alone_by_cob <-
  read_excel(
    path = exported_file,
    sheet = "D",
    range = "D26:K26",
    col_names = living_alone_table_header[-1],  # remove the first variable in the constructed column names from above since we're not using the urban rural column here
  ) %>%
  pivot_longer(cols = everything(), names_to = c("living_alone","country_of_birth"), names_pattern = "^(alone|total)\\_(.*)$", values_to = "n")  # split the columns containing multiple variables (living alone or not & CoB)

## test that we get the right numbers
living_alone_by_cob_and_urban_rural %>%
  group_by(living_alone, country_of_birth) %>%
  summarise(n_sum = sum(n), .groups="drop") %>%
  left_join(living_alone_by_cob)
```

## `M1` to `M3`

These are logistic models:

* M1 is a stratified regression (stratified on 3 urban-rural categories, meaning three separate models) of living alone on the other variables: time between census and death, marital status, age at death, country of birth, tenure (housing), and sex
* M2 is the same model as M1 but the dependent variable is dying in an institution
* M3 is like M1 but the 3 categories of urban-rurality are included as an independent variable

As far as I can tell, for the purposes of the paper, M1 and M2 were used, and only the Rural (remote and accessible) stratum was considered (interpreted as the destination of migrations).

There are univariate models in the bottom of each sheet.

### M1

```{r}
regression_columns <- c("b", "se", "Wald", "df", "p_value", "or")  # copied from excel & adjusted for more r-like labelling

m1_regression_variables <- c(
  "country_of_birth",
  "age_at_death",
  "sex_at_death",
  "housing_tenure",
  "time_to_death",
  "marital_status"
)

m1_univariate_regression_ranges <- c(
  "C125:I128", # CoB  - this is lsited in the first step of the regressio nrather than unadjusted regressions later!
  "C285:I288",  # age
  "C300:I301",  # sex
  "C314:I317",  # housing tenure
  "C330:I337",  # time to death
  "C348:I352"  # marital status
)


m1_remote_univariate_regressions <-
  map2(
    .x = m1_regression_variables,
    .y = m1_univariate_regression_ranges,
    .f = 
      ~read_excel(
      path = exported_file,
      sheet = "M1",
      range = .y,
      col_names = c(.x, regression_columns),  # remove the first variable in the constructed column names from above since we're not using the urban rural column here
      col_types = c("text",rep("numeric",times=6))
    )
  ) %>% set_names(m1_regression_variables)

## The country of birth value labels are missing so we need to change them manually!
regression_country_of_birth_categories <- country_of_birth_categories[-4]
regression_country_of_birth_categories[1] <- "Scotland (ref: Scotland)"
m1_remote_univariate_regressions$country_of_birth$country_of_birth[1:3] <- regression_country_of_birth_categories

## The sex variable is binary and has only one entry - to make it have the same format as the others we add another row for the reference column
m1_remote_univariate_regressions$sex_at_death <-
  m1_remote_univariate_regressions$sex_at_death %>%
  add_case(sex_at_death = m1_remote_univariate_regressions$sex_at_death$sex_at_death[1], .before=1)
m1_remote_univariate_regressions$sex_at_death$sex_at_death[2] <- "male"
  

m1_multiple_regression_ranges <- c(
  "C239:I241",  # CoB
  "C242:I244",  # age
  "C245:I245",  # sex 
  "C246:I248",  # housing tenure
  "C249:I255",  # time to death
  "C256:I259",  # marital status
  "C260:I260"  #  constant
)

m1_remote_multiple_regressions <-
  map2(
    .x = c(m1_regression_variables,"constant"),
    .y = m1_multiple_regression_ranges,
    .f = 
      ~read_excel(
      path = exported_file,
      sheet = "M1",
      range = .y,
      col_names = c(.x, regression_columns),  # remove the first variable in the constructed column names from above since we're not using the urban rural column here
      col_types = c("text",rep("numeric",times=6))
    )
  ) %>% set_names(c(m1_regression_variables,"constant"))

## the variable labels are missing in the multiple regression, but they can be copied over from the univariate regressions!
m1_remote_multiple_regressions$country_of_birth$country_of_birth <- m1_remote_univariate_regressions$country_of_birth$country_of_birth[1:3]
m1_remote_multiple_regressions$age_at_death$age_at_death <- m1_remote_univariate_regressions$age_at_death$age_at_death[1:3]
m1_remote_multiple_regressions$sex_at_death <-  # we'll create an extra row for the sex variable which is binary & didn't have a separate reference entry like all the others
  m1_remote_multiple_regressions$sex_at_death %>%
  add_case(sex_at_death=NA, .before = 1)
m1_remote_multiple_regressions$sex_at_death$sex_at_death <- m1_remote_univariate_regressions$sex_at_death$sex_at_death[1:2]
m1_remote_multiple_regressions$housing_tenure$housing_tenure <- m1_remote_univariate_regressions$housing_tenure$housing_tenure[1:3]
m1_remote_multiple_regressions$time_to_death$time_to_death <- m1_remote_univariate_regressions$time_to_death$time_to_death[1:7]
m1_remote_multiple_regressions$marital_status$marital_status <- m1_remote_univariate_regressions$marital_status$marital_status[1:4]
```


### M2

```{r}
m2_regression_variables <- c(
  "country_of_birth",
  "age_at_death",
  "sex_at_death",
  "housing_tenure",
  "time_to_death",
  "living_alone",
  "marital_status"
)

m2_univariate_regression_ranges <- c(
  "C117:I120", # CoB  - this is lsited in the first step of the regressio nrather than unadjusted regressions later!
  "C322:I325",  # age
  "C336:I337",  # sex
  "C348:I351",  # housing tenure
  "C362:I369",  # time to death
  "C380:I381",  # living alone
  "C392:I396"  # marital status
)


m2_remote_univariate_regressions <-
  map2(
    .x = m2_regression_variables,
    .y = m2_univariate_regression_ranges,
    .f = 
      ~read_excel(
      path = exported_file,
      sheet = "M2",
      range = .y,
      col_names = c(.x, regression_columns),  # remove the first variable in the constructed column names from above since we're not using the urban rural column here
      col_types = c("text",rep("numeric",times=6))
    )
  ) %>% set_names(m2_regression_variables)

## living alone and sex are binary, for both we should set a reference level so they are in the same format as the other variables
m2_remote_univariate_regressions$sex_at_death <-
  m2_remote_univariate_regressions$sex_at_death %>%
  add_case(sex_at_death = m2_remote_univariate_regressions$sex_at_death$sex_at_death[1], .before=1)
  
m2_remote_univariate_regressions$living_alone <-
  m2_remote_univariate_regressions$living_alone %>%
  add_case(living_alone = m2_remote_univariate_regressions$living_alone$living_alone[1], .before=1)
m2_remote_univariate_regressions$living_alone$living_alone[1] <- "alone (ref: alone)"
m2_remote_univariate_regressions$living_alone$living_alone[2] <- "with others"

## all value labels are missing here so we can copy them from M1, except for living alone
for (variable in names(m1_remote_univariate_regressions)) {
  m2_remote_univariate_regressions[[variable]][[variable]] <- m1_remote_univariate_regressions[[variable]][[variable]]
}

m2_multiple_regression_ranges <- c(
  "C277:I279",  # CoB
  "C280:I282",  # age
  "C283:I283",  # sex 
  "C284:I286",  # housing tenure
  "C287:I293",  # time to death
  "C294:I294",  # living alone
  "C295:I298",  # living alone
  "C299:I299"  #  constant
)

m2_remote_multiple_regressions <-
  map2(
    .x = c(m2_regression_variables,"constant"),  # we include the constant term here also
    .y = m2_multiple_regression_ranges,
    .f = 
      ~read_excel(
      path = exported_file,
      sheet = "M2",
      range = .y,
      col_names = c(.x, regression_columns),  # remove the first variable in the constructed column names from above since we're not using the urban rural column here
      col_types = c("text",rep("numeric",times=6))
    )
  ) %>% set_names(c(m2_regression_variables,"constant"))


## living alone and sex are binary, for both we should set a reference level so they are in the same format as the other variables
m2_remote_multiple_regressions$sex_at_death <-
  m2_remote_multiple_regressions$sex_at_death %>%
  add_case(sex_at_death = m2_remote_multiple_regressions$sex_at_death$sex_at_death[1], .before=1)
  
m2_remote_multiple_regressions$living_alone <-
  m2_remote_multiple_regressions$living_alone %>%
  add_case(living_alone = m2_remote_univariate_regressions$living_alone$living_alone[1], .before=1)

## the variable labels are missing in the multiple regression, but they can be copied over from the univariate regressions!
for (variable in names(m2_remote_univariate_regressions)) {
  number_entries <- length(m2_remote_multiple_regressions[[variable]][[variable]])
  m2_remote_multiple_regressions[[variable]][[variable]] <-
    m2_remote_univariate_regressions[[variable]][[variable]][1:number_entries]
}
```


# Notes on data

* Individuals with missing marital status were excluded due to very small N;
* Logistic models used complete data only


# Processed tables

```{r}
tables <- list()

tables$m1_remote_unadjusted_odds <-
  map2_df(
    .x = m1_remote_univariate_regressions,
    .y = names(m1_remote_univariate_regressions),
    .f = function(x,y) {
      tbl <- select(x, 1, or) %>%
      mutate(variable = y) %>%
      rename(level=1) %>%
      filter(level!="Constant") %>%
      relocate(variable, level, or)
      
      tbl$level[1] <- str_replace(string = tbl$level[1], pattern=".*ref\\:\\s(.*)\\)", replacement="Reference: \\1")  # rename the reference level
      # tbl$or[1] <- 1L  # set reference OR to 1
      tbl$or[1] <- NA  # set reference OR to NA
      return(tbl)
    } 
  )

tables$m1_remote_adjusted_odds <-
  map2_df(
    .x = m1_remote_multiple_regressions[names(m1_remote_multiple_regressions)!="constant"],  # don't incldue the Constant term
    .y = names(m1_remote_univariate_regressions[names(m1_remote_multiple_regressions)!="constant"]),  # see above comment
    .f = function(x,y) {
      tbl <- select(x, 1, or) %>%
      mutate(variable = y) %>%
      rename(level=1) %>%
      filter(level!="Constant") %>%
      relocate(variable, level, or)
      
      tbl$level[1] <- str_replace(string = tbl$level[1], pattern=".*ref\\:\\s(.*)\\)", replacement="Reference: \\1")  # rename the reference level
      # tbl$or[1] <- 1L  # set reference OR to 1
      tbl$or[1] <- NA  # set reference OR to NA
      return(tbl)
    } 
  )

tables$m1_odds <-
  full_join(
    x = tables$m1_remote_unadjusted_odds %>% rename(unadjusted_or = or),
    y = tables$m1_remote_adjusted_odds %>% rename(adjusted_or = or),
    by = c("variable", "level")
  )


tables$m2_remote_unadjusted_odds <-
  map2_df(
    .x = m2_remote_univariate_regressions,
    .y = names(m2_remote_univariate_regressions),
    .f = function(x,y) {
      tbl <- select(x, 1, or) %>%
      mutate(variable = y) %>%
      rename(level=1) %>%
      filter(level!="Constant") %>%
      relocate(variable, level, or)
      
      tbl$level[1] <- str_replace(string = tbl$level[1], pattern=".*ref\\:\\s(.*)\\)", replacement="Reference: \\1")  # rename the reference level
      # tbl$or[1] <- 1L  # set reference OR to 1
      tbl$or[1] <- NA  # set reference OR to NA
      return(tbl)
    } 
  )

tables$m2_remote_adjusted_odds <-
  map2_df(
    .x = m2_remote_multiple_regressions[names(m2_remote_multiple_regressions)!="constant"],  # don't incldue the Constant term
    .y = names(m2_remote_univariate_regressions[names(m2_remote_multiple_regressions)!="constant"]),  # see above comment
    .f = function(x,y) {
      tbl <- select(x, 1, or) %>%
      mutate(variable = y) %>%
      rename(level=1) %>%
      filter(level!="Constant") %>%
      relocate(variable, level, or)
      
      tbl$level[1] <- str_replace(string = tbl$level[1], pattern=".*ref\\:\\s(.*)\\)", replacement="Reference: \\1")  # rename the reference level
      # tbl$or[1] <- 1L  # set reference OR to 1
      tbl$or[1] <- NA  # set reference OR to NA
      return(tbl)
    } 
  )

tables$m2_odds <-
  full_join(
    x = tables$m2_remote_unadjusted_odds %>% rename(unadjusted_or = or),
    y = tables$m2_remote_adjusted_odds %>% rename(adjusted_or = or),
    by = c("variable", "level")
  )
# TODO: look at formatting tables with `gt`
```

# Processed figures

```{r}
figures <- list()

# TODO: complete below!
figures$alone_by_country_of_birth_and_urban_rural <- ggplot(
  data = living_alone_by_cob_and_urban_rural %>% pivot_wider(names_from=living_alone, values_from=n) %>% mutate(proportion=alone/total) %>% filter(country_of_birth!="All CoB")
) +
  geom_col(aes(x=urban_rural_6_categories, y=proportion, group=country_of_birth, fill=country_of_birth), position = position_dodge()) +
  NULL
```

